Allocation Removal by Partial Evaluation in a Tracing JIT

### 2010-11-18 talk/pepm2011/paper.tex
try to use the word "hot" only after it is defined

% XXX insert "practical" somewhere

@@

A recently popular approach to implementing just-in-time compilers for dynamic
languages is that of a tracing JIT. A tracing JIT works by observing the running
program and recording its hot spots into \emph{linear execution traces}. Those
traces are optimized and turned into machine code.

@@

\begin{enumerate}
    \item A description of an efficient and effective algorithm for removing
          object allocations in a tracing JIT.
    \item A characterization of this algorithm as partial evaluation.
    \item Performance benchmarks for this algorithm.
\end{enumerate}

@@

The core idea of tracing JITs is to focus the optimization effort of the JIT
compiler on the commonly executed, \ie \emph{hot} paths of the core loops of the program and to just use an
interpreter for the less commonly executed parts. VMs that use a tracing JIT are
mostly mixed-mode execution environments, they contain both an interpreter and a
JIT compiler. By default the interpreter is used to execute the program, doing
some light-weight profiling at the same time. This profiling is used to identify
the hot loops\reva{what is that? clarify} of the program. If a hot loop is found in that way, the
interpreter enters a special \emph{tracing mode}. In this tracing mode, the
interpreter tries to record all operations that it is executing while running one
iteration of the hot loop. This history of executed operations of one loop is
called a \emph{trace}. Because the trace corresponds to one iteration of a loop,
it always ends with a jump to its own beginning. The trace also contains all
operations that are performed in functions that were called in the loop, thus a
tracing JIT automatically performs inlining.
This trace of operations subsequently forms the basis of the generated code. The trace is
first optimized, and then turned into machine code. Both optimization
and machine code generation are simple, because the traces are linear. This
linearity makes many optimizations a lot more tractable, and the inlining that
happens gives the optimizations automatically more context to work with.

>>>>

@@

A recently popular approach to implementing just-in-time compilers for dynamic
languages is that of a tracing JIT. A tracing JIT works by observing the running
program and recording its commonly executed parts into \emph{linear execution traces}. Those
traces are optimized and turned into machine code.

@@

\begin{enumerate}
    \item A description of a practical, efficient and effective algorithm for
    removing object allocations in a tracing JIT.
    \item A characterization of this algorithm as partial evaluation.
    \item Performance benchmarks for this algorithm.
\end{enumerate}

@@

The core idea of tracing JITs is to focus the optimization effort of the JIT
compiler on the commonly executed, \ie \emph{hot} paths of the core loops of the program and to just use an
interpreter for the less commonly executed parts. VMs that use a tracing JIT are
mostly mixed-mode execution environments, they contain both an interpreter and a
JIT compiler. By default the interpreter is used to execute the program, doing
some light-weight profiling at the same time. This profiling is used to identify
the hot loops of the program. If a hot loop is found in that way, the
interpreter enters a special \emph{tracing mode}. In this tracing mode, the
interpreter tries to record all operations that it is executing while running one
iteration of the hot loop. This history of executed operations of one loop is
called a \emph{trace}. Because the trace corresponds to one iteration of a loop,
it always ends with a jump to its own beginning. The trace also contains all
operations that are performed in functions that were called in the loop, thus a
tracing JIT automatically performs inlining.
This trace of operations subsequently forms the basis of the generated code. The trace is
first optimized, and then turned into machine code. Both optimization
and machine code generation are simple, because the traces are linear. This
linearity makes many optimizations a lot more tractable, and the inlining that
happens gives the optimizations automatically more context to work with.
